---
title: "Import sleep diaries"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

## Preface

This document imports the `sleepdiaries` and shows descriptive statistics for the site.

## Setup

```{r}
#| message: false
library(tidyverse)
library(LightLogR)
library(glue)
library(readxl)
library(gt)
library(gtsummary)

remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c("labeling",
  "radio_factors",
  "time_summaries",
  "prepare_codebook",
  "filefinder",
  "general_parameters",
  "coltype_checker",
  "tables"
) |> walk(\(x) source(paste0(remote, x, ".R")))
```

Collect the necessary external information

```{r}
codebook <- 
  read_csv(
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/codebook/MeLiDosMorningSleepDiaries_DataDictionary_2024-10-16.csv", show_col_types = FALSE
  )
#clean up labels
codebook <-
  codebook |> 
    mutate(
      `Field Label` = 
        `Field Label` |> 
        str_remove_all("<div class=\"rich-text-field-label\">|<p>|<em>|</em>|</p>|</div>")
    )
```


## Collect files

The following files contain sleep-diary information.

```{r}
#path to participants
path_part1 <- "../data/raw/individual"
#path to questionnaire
path_part2 <- "/continuous/sleepdiary"
#getting all subfolders
folders <- dir(path_part1)
#creating complete folder names
paths <- glue("{path_part1}/{folders}{path_part2}")
#collecting file names
files <- list.files(paths, full.names = TRUE)
```

## Import files

```{r}
sleepdiary <-
  read_csv2(files, show_col_types = FALSE) |> 
  drop_na(redcap_repeat_instance) |> 
  mutate(
    across(
      c(sleep, bedtime, offset, out_ofbed), 
      \(x) parse_date_time(x, orders = "dmyHM", tz = "UTC")
      ),
    record_id = paste0("MPI_S", record_id),
    across(
      c(awakenings, awake_duration),
      parse_number
    ),
    sleepquality = 
      sleepquality___1 + 2*sleepquality___2 + 
      3*sleepquality___3 + 4*sleepquality___4 + 5*sleepquality___5,
    daytype2 = daytype2___1 + 2*daytype2___2
    )
```

### Check column types

```{r}
should_POSIXct <- c("bedtime", "sleep", "offset", "out_ofbed")
should_numeric <- c("sleepdelay", "awakenings", "awake_duration", "sleepquality", "daytype2")
should_character <- c("record_id", "comments")

stopifnot(
  "all of should_POSIXct need to be part of the dataset and of type POSIXct" =
  should_POSIXct %in% (sleepdiary |> select(where(is.POSIXct)) |> names()),
  "all of should_numeric need to be part of the dataset and of type numeric" =
  should_numeric %in% (sleepdiary |> select(where(is.numeric)) |> names()),
  "all of should_character need to be part of the dataset and of type character" =
  should_character %in% (sleepdiary |> select(where(is.character)) |> names())
  )
```

### Select relevant variables &nonempty rows

```{r}
sleepdiary <- 
sleepdiary |> 
  select(all_of(c(should_character, should_POSIXct, should_numeric)))
```


### Set sleep time zone
```{r}
site <-  "MPI"
sleepdiary <- 
  sleepdiary |> 
  mutate(across(where(is.POSIXct), 
                \(x) force_tz(x, tzs[[site]]))
         )
```


### Label variables

```{r}
sleepdiary <-
sleepdiary |> 
  add_radio_factors(codebook, 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook, var_col = `Variable / Field Name`, label_col = `Field Label`)
```

### Calculate sleep and sleepduration

```{r}
sleepdiary <- 
sleepdiary |> 
  rename(wake = offset, sleepprep = sleep) |> 
  mutate(sleep = sleepprep + dminutes(sleepdelay),
         sleep_duration = (wake - sleep)/60
  )

attr(sleepdiary$sleep, "label") <- "Sleep onset (calculated)"
attr(sleepdiary$sleep_duration, "label") <- "Sleep duration (calculated)"
attr(sleepdiary$sleep_duration, "units") <- "hrs"

```


### Ensure nighttime awake duration when there are no awakenings

```{r}
sleepdiary <-
  sleepdiary |> 
  mutate(
    awake_duration = replace_when(awake_duration,
                                  awakenings == 0 ~ 0)
  )
```

## Summarize diary results

```{r}
table_sleep <- 
sleepdiary |> 
  tbl_summary(include = -c(comments, record_id),
              statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", 
                               all_categorical() ~ "{n} ({p}%)",
                               c(bedtime, sleep, sleepprep)  ~ "{time_median} ({nighttime_p25}, {nighttime_p75})",
                               c(wake, out_ofbed)  ~ "{time_median} ({daytime_p25}, {daytime_p75})"
                               ),
              type = awakenings ~ "continuous",
              missing_text = "missing") |> 
  add_n() |> 
  bold_labels() |> 
  modify_header(label = "**Morning sleep diary**") |> 
  modify_footnote_header("Nighttime variables center on midnight, daytime variables on noon; median for time is based on circular time", columns = stat_0, replace = FALSE)

table_sleep

gtsave(table_sleep |> as_gt(), filename = "../output/tables/table_sleepdiary.png")
```

### Translate comments

Native language is translated into English with AI and later checked by a site researcher.

```{r}
#| label: translate into english with AI
# 
# library(ellmer)
# 
# #Providing the relevant codebook portions
# codebook_red <-
# codebook|>
#   pmap(~ paste(paste(names(codebook), c(...), sep = ": "), collapse = ", ")) |>
#   list_c() |>
#   paste0(collapse = "newline: ")
# 
# chat <- chat_openai(paste0("Copy content and make translations according to specific instructions. They represent answers to questionnaires in a scientific field study."))
# 
# #Providing the input
# data_red <-
# sleepdiary|>
#   select(record_id, comments)
# data_red <-
#   data_red |>
#   pmap(~ paste(paste(names(data_red), c(...), sep = ": "), collapse = ", "))
# 
# #creating an output structure
# type_data <- type_object(
#   record_id = type_string("use the record_id information"),
#   comments = type_string("copy the original information here", required = FALSE),
#   comments_english = type_string("translation of the original information into english", required = FALSE)
# )
# 
# data_llm <-
# parallel_chat_structured(
#   chat,
#   data_red,
#   type = type_data,
#   rpm = 500,
#   max_active = 100
# )
# 
# #Ensure that no NA is caught as string
# data_llm <-
#   data_llm |>
#   mutate(comments = case_when(comments == "NA" ~ NA, .default = comments),
#          comments_english = case_when(comments_english == "NA" ~ NA, .default = comments_english))
# 
# #check that input and output are identical
# stopifnot("Input must by identical to output check" =
#             all(data_llm$comments == sleepdiary$comments, na.rm = TRUE))
# stopifnot("Input must by identical to output check" =
#             all(data_llm$record_id == sleepdiary$record_id, na.rm = TRUE))
# 
# data_llm <-
#   data_llm |>
#   distinct(record_id, comments, .keep_all = TRUE)
# 
# path <- "../data/AI_translations/"
# if(!dir.exists(path)) dir.create(path, recursive = TRUE)
# write_csv(data_llm, "../data/AI_translations/sleepdiary.csv")

```

```{r}
data_llm <- 
  read_csv("../data/AI_translations/sleepdiary.csv")
#add output to original
sleepdiary <- sleepdiary |> left_join(data_llm, by = c("record_id", "comments"))
attr(sleepdiary$comments_english, "label") <- "Comments (English translation)"
```

### Sort by date

```{r}
sleepdiary <- 
  sleepdiary |> 
  arrange(record_id, bedtime)
```

### Export

```{r}
sleepdiary <- sleepdiary |> rename(Id = record_id)
path <- "../data/imported/continuous/"
if(!dir.exists(path)) dir.create(path, recursive = TRUE)
save(sleepdiary, file = "../data/imported/continuous/sleepdiaries.RData")
```

